---
title: "COVID-19 in Newfoundland and Labrador"
author: "Amy Hurford, Zahra Mohammadi, and Francis Anokye"
output:
  html_document: default
  word_document: default
  pdf_document: default
fontsize: 11pt
---
[Memorial University of Newfoundland and Labrador,](https://www.mun.ca/math/)
[University of Guelph,](https://mathstat.uoguelph.ca/)
[Canadian Network for Modelling Infectious Disease,](https://canmod.net/)
[Mathematics for Public Health,](http://www.fields.utoronto.ca/activities/public-health)
[One Health Modelling Network for Emerging Infections](https://www.yorku.ca/science/cdm/2021/04/09/ccdm-network-to-model-emerging-infectious-diseases-receives-2-5-million-in-federal-funding/)

```{r setup, include=FALSE}
require(ggplot2)
require(patchwork)
data <- read.csv('https://raw.githubusercontent.com/ahurford/covid-nl/master/NL.csv', fill=TRUE)

test <- read.csv('https://raw.githubusercontent.com/ahurford/covid-nl/master/Test_NL(twosources).csv', fill=TRUE)

## These data points are only reported on twitter: Health and community services NL
# Dec 28 4174
# Jan 1 4680
# Jan 2 3354
# Jan 8 4041

## Unfortunately quite a lot of this script is not automated and requires manual entry.
# 1. Manually enter data in NL.csv and two-sources.csv.
#  - From the PSAs. For each RHA you need to add the number of male + female + unknown cases.
#  - For Eastern health subtract the number of tests that were send out of province.
#  - Active hospitalizations & deaths are reported after the regional breakdown. Deaths is cumulative, so if
#   1 death is reported add +1 to the last entry. For hosptializations it is not cumulative - just the number   #    they report as in hospital on that day.
#    The number of cases being sent out of provinces needs to be manually entered in this code.

# 2. Knit script.
# 3. Check doubling times, i.e. DT.NL. The forth entry is significance. If >0.05 report as "Plateau". If significant report as doubling time or halving time - whichever function is called HT() or DT(). If the first element of DT.NL is negative switch from DT() to HT(). Check all of DT.eastern, DT.central, DT.western, DT.labrador, DT.NL, and write in the correct values in the text.
# 4. Check the testing regression. Is a2<a1? In the text we say it is.
# 5. Check the hospitalization and mortality graphs. If the calibration is poor, choose new values of NLh and NLm
# 6. Manually move and rename omicron-data.html to index.html in the docs folder

## This averages the test results across days when there is no reporting
test1 = test$total.test
test1[2] = mean(c(test1[1],test1[3]))
test1[4] = mean(c(test1[3],test1[5]))
test1[11] = mean(c(test1[10],test1[14])-test1[10])*(1/4)+test1[10]
test1[12] = mean(c(test1[10],test1[14])-test1[10])*(2/4)+test1[10]
test1[13] = mean(c(test1[10],test1[14])-test1[10])*(3/4)+test1[10]
test1 = diff(test1)

## This averages the data across days when there is no reporting
data[1,2:7] = data[1,2:7]/2
# Need to add in December 16
new1 = data[2,]
L = length(data[,1])
new1[1,1] = "2021-12-16"
new1[1,2:7] = new1[2:7]/2
data = rbind(data[1,],new1,data[2:L,])
data[3,2:7] = data[2,2:7]
data[11,2:7] = data[11,2:7]/3
L = length(data[,1])
new1[1,1] = "2021-12-25"
new1[1,2:8] = data[11,2:8]
new2 = new1
new2[1,1] = "2021-12-26"
data = rbind(data[1:10,],new1, new2,data[11:L,])

## This is the regression on tests
L=length(data$total)
cases = data$total[2:L]
# Remove Christmas dates
test1 = test1[-(10:12)]
dd = tail(data$date,-1)
dd = dd[-(10:12)]
cases = cases[-(10:12)]

# Lag cases by 3 days:
lag = 3
test3 = tail(test1,-lag)
dd3 = head(dd,-lag)
cases3 = head(cases,-lag)
L = length(cases3)

# Break Jan 3 when changes to PCR were announced
ind = which(dd3 == "2022-01-03")
ind2 = which(dd3 == "2022-01-07") # Excludes Jan 10 which was abnormally high day for tests > 7000.
# The above will need to be modified if testing to notification lag is different
cases1 = c(cases3[1:ind], rep(0,(L-ind)))
cases2 = c(rep(0,ind), cases3[(ind+1):L])
popn = rep((317664+91201+76608+36069),L)
popn = popn[-(ind2:L)]
test3 = test3[-(ind2:L)]
cases2 = cases2[-(ind2:L)]
cases1 = cases1[-(ind2:L)]
mod = lm(test3 ~ cases1 + cases2 + popn + 0)

a1 = round(unname(coef(mod)[1]),2)
a2 = round(unname(coef(mod)[2]),2)
b = round(unname(coef(mod)[3]),2)

## Making a dataframe for plotting
NL.data = data
L = length(data[,1])
data2 = data.frame(date = rep(data$date,4), cases = c(data$EH, data$CH, data$WH, data$LG), RHA = rep(c("Eastern", "Central", "Western", "Labrador-Grenfell"), each=L)) 
data3 = data.frame(date = rep(data$date,5), cases = 100000*c(data$EH/317664, data$CH/91201, data$WH/76608, data$LG/36069, data$total/(317664+91201+76608+36069)), RHA = rep(c("Eastern", "Central", "Western", "Labrador-Grenfell", "Newfoundland & Labrador"), each=L))

## Correcting the data for the change in PCR testing on Jan 3 & distributing the test results for the
## out-of-province reporting across a range of days.
corr = a1-a2
ind = min(which(data3$date=="2022-01-06"))
# PCR tsting change to EH
EH.corr = data$EH + corr*c(rep(0,(ind-1)),data$EH[(ind-3):(L-3)])
# Re-distributing the cases reported on Jan 10, which were collected Dec 29 to 
i1=min(which(data3$date == "2021-12-29"))
EH.corr[i1:ind] = EH.corr[i1:ind]+EH.corr[i1:ind]/sum(EH.corr[i1:ind])*680
#323 on Dec 11
EH.corr[(i1+1):(ind+1)] = EH.corr[(i1+1):(ind+1)]+EH.corr[(i1+1):(ind+1)]/sum(EH.corr[(i1+1):(ind+1)])*323
#229 on Dec 12
EH.corr[(i1+2):(ind+2)] = EH.corr[(i1+2):(ind+2)]+EH.corr[(i1+2):(ind+2)]/sum(EH.corr[(i1+2):(ind+2)])*229
#166 on Dec 13
EH.corr[(i1+3):(ind+3)] = EH.corr[(i1+3):(ind+3)]+EH.corr[(i1+3):(ind+3)]/sum(EH.corr[(i1+3):(ind+3)])*166
# 71 on Jan 14
EH.corr[(i1+4):(ind+4)] = EH.corr[(i1+4):(ind+4)]+EH.corr[(i1+4):(ind+4)]/sum(EH.corr[(i1+4):(ind+4)])*71
# 71 on Jan 15
EH.corr[(i1+5):(ind+5)] = EH.corr[(i1+5):(ind+5)]+EH.corr[(i1+5):(ind+5)]/sum(EH.corr[(i1+5):(ind+5)])*3

# Corrections for PCR changes
CH.corr = data$CH + corr*c(rep(0,(ind-1)),data$CH[(ind-3):(L-3)])
WH.corr = data$WH + corr*c(rep(0,(ind-1)),data$WH[(ind-3):(L-3)])
LG.corr = data$LG + corr*c(rep(0,(ind-1)),data$LG[(ind-3):(L-3)])
NL.corr = EH.corr + CH.corr + WH.corr + LG.corr

# Corrected data for plotting
data2.corr = data.frame(date = data$date, EH = EH.corr, NL = NL.corr, WH= WH.corr, LG = LG.corr, CH = CH.corr) 
data3.corr = data.frame(date = rep(data$date,5), cases = 100000*c(EH.corr/317664, CH.corr/91201, WH.corr/76608, LG.corr/36069,(EH.corr+CH.corr+WH.corr+LG.corr)/(317664+91201+76608+36069)), RHA = rep(c("Eastern", "Central", "Western", "Labrador-Grenfell", "Newfoundland & Labrador"), each=L))


# Doubling time and halving time functions. If the first value of the result is negative, you need to change
# from calling HT() or DT() to the other one.

# Window of days over which to calculate the DT or HT.
T<-7
DT = function(){
mod = nls(val ~ a*exp(r*seq(1,T)), 
               start = list(a = 100, r = 0.2))
# The code below was slow, so I commented it out.
#CI = confint(mod, level=0.95)
#CI.lower = max(CI[2,1],0)
#CI.upper = CI[2,2]
lambda = unname(coef(mod)[2])
DT = c(round(log(2)/lambda,2), NA, NA)
p1=summary(mod)$coefficients[,4]
p = unname(p1[2])
return(c(DT,p))
}

HT = function(){
mod = nls(val ~ a*exp(-r*seq(1,T)), 
               start = list(a = 300, r = 0.2))
## The code below was slow so I commented it out
#CI = confint(mod, level=0.95)
#CI.lower = max(CI[2,1],0)
#CI.upper = CI[2,2]
lambda = unname(coef(mod)[2])
DT = c(-round(log(.5)/lambda,2), NA, NA)
p1=summary(mod)$coefficients[,4]
p = unname(p1[2])
return(c(DT,p))
}

# For EH it is quite likely there are tests that have been sent out-of-province
# that have yet to be redistributed so I don't consider the last 3 days of data.
val <- head(tail(EH.corr,(T+3)),-3)
DT.eastern = HT()
print(DT.eastern)
val <- tail(CH.corr,T)
DT.central = HT()
val <- tail(WH.corr,T)
DT.western = HT()
val <- tail(LG.corr,T)
DT.labrador = HT()
# See comments for EH
val <- head(tail(NL.corr,(T+3)),-3)
DT.NL = HT()
```

<div class="alert alert-info">

On this website we:

- visualize COVID-19 cases reported by [Newfoundland and Labrador (NL)](https://www.gov.nl.ca/releases/covid-19-news/);

- estimate the doubling time of reported cases; and

- predict hospitalizations and deaths.

The data begins on December 15, 2021, when the [first Omicron case](http://ntv.ca/province-confirms-first-case-of-omicron-variant-introduces-new-public-health-measures/) was reported in NL.

</div>

## Data limitations
On January 12, Dr. Rahman presented the provincial modelling. Some of Dr. Rahman's graphs considered [specimen date](https://twitter.com/PeterCBC/status/1481323100163784708) rather than report date. These data, by specimen date, are a more accurate representation of the COVID-19 outbreak in NL. This website considers only report date as specimen dates are not reported in the [Public Service Advisories]((https://www.gov.nl.ca/releases/covid-19-news/)).

## Testing

[Recommendations for PCR testing](https://www.gov.nl.ca/releases/2022/health/0103n02/) changed in NL on January 3rd, 2022, such that anyone that is a close contact of a case and has symptoms of COVID-19, should consider themselves a positive case and not schedule a PCR test. In British Columbia (BC), people under 65 are discouraged from seeking PCR testing. The [BC Modelling Group](https://bccovid-19group.ca/post/2022-01-06-report/COVID19group-2022-01-06.pdf) estimates that the number of cases reported in BC would be 3.6 times higher without this change in testing.

To assess the impact of the change in PCR testing recommendations, we assumed that the number of tests reported for NL on a given day is the sum of two components:

$a_1$ or $a_2$: the number of tests arising due to a contact with a COVID-19 case reported 3 days earlier, and
$b$: the number of tests performed per person in NL;

where $a_1$ is prior to January 6, and $a_2$ is after January 6th. We assume a 3 day lag between testing and reporting, such that if on January 3 symptomatic contacts of a case did not get tested the effect would first appear in the reporting on January 6. Using the data from the [Public Service Annoucements](https://www.gov.nl.ca/releases/covid-19-news/) we estimate that:

$a_1=$  `r a1`  tests per case 3 days prior (to January 3)

$a_2=$ `r a2` tests per case 3 days prior (after January 3)

$b=$ `r b` tests per person in NL.

This suggests that after January 3rd, `r a1 - a2` fewer tests were performed per case because symptomatic contacts of cases did not get tested.

### Adjusted cases: correcting for the PCR testing changes and out-of-province test results

To correct for the change in PCR testing recommendations (beginning January 3), we assume that beginning January 6th the reporting reflects `r a1-a2` fewer tests per case, and that all these tests that did not occur would have been positive. We use this adjustment so that reported cases can be compared before and after January 6. 

[On January 10, 2022, NL reported 680 cases](https://www.gov.nl.ca/releases/2022/health/0110n04/) arising from tests sent out of province December 29 to January 6th. The numbers from the [Public Service Announcement](https://www.gov.nl.ca/releases/2022/health/0110n04/) suggest that all 680 cases were in Eastern Health. To correct the case data these 680 cases are distributed in Eastern Health in proportion to the number of reported cases on each day December 29 to January 6th. The addition of these cases is after the correction for the PCR testing change.

Cases from out-of-province testing have also been reported on the days after January 10. These cases are redistributed similarly, where the January 11 cases are distributed from December 30 to January 7th, and January 12th cases are distributed from December 31 to January 8th, and so forth.

All subsequent calculations are performed on the adjusted data.

## Adjusted cases

```{r, echo=FALSE, warning = FALSE}
g1=ggplot(data3.corr, aes(x=date, y=cases, group = RHA)) +
geom_line(aes(color=RHA))+
  geom_point(aes(color=RHA))+
  xlab("")+
  ylab("")+
  ggtitle("Adjusted cases per 100,000 people in each NL Regional Health Authority")+
  #ggtitle(paste("At entry: ", exp.on.entry,"days since exposure"))+
theme_classic()+
  theme(axis.text.x = element_text(angle = 90),legend.title = element_blank())

g=ggplot() +
geom_line(data=data2.corr, aes(x=as.Date(date), y=NL, group=1))+
geom_point(aes(x=as.Date(data$date), y=data$EH+data$WH+data$CH+data$LG))+
  xlab("")+
  ylab("Reported and adjusted cases")+
  ggtitle("Newfoundland and Labrador")+
  #ggtitle(paste("At entry: ", exp.on.entry,"days since exposure"))+
theme_classic()+
  scale_x_date(date_breaks = "1 day", date_labels = "%d %b")+
  theme(axis.text.x = element_text(angle = 90),legend.title = element_blank())

g.EH=ggplot() +
geom_line(data=data2.corr, aes(x=as.Date(date), y=EH, group=1))+
geom_point(aes(x=as.Date(data$date), y=data$EH))+
  xlab("")+
  ylab("Reported and adjusted cases")+
  ggtitle("Eastern Health")+
  #ggtitle(paste("At entry: ", exp.on.entry,"days since exposure"))+
theme_classic()+
  scale_x_date(date_breaks = "2 day", date_labels = "%d %b")+
  theme(axis.text.x = element_text(angle = 90),legend.title = element_blank())

g.CH=ggplot() +
geom_line(data=data2.corr, aes(x=as.Date(date), y=CH, group=1))+
geom_point(aes(x=as.Date(data$date), y=data$CH))+
  xlab("")+
  ylab("Reported and adjusted cases")+
  ggtitle("Central Health")+
  #ggtitle(paste("At entry: ", exp.on.entry,"days since exposure"))+
theme_classic()+
  scale_x_date(date_breaks = "2 day", date_labels = "%d %b")+
  theme(axis.text.x = element_text(angle = 90),legend.title = element_blank())

g.WH=ggplot() +
geom_line(data=data2.corr, aes(x=as.Date(date), y=WH, group=1))+
geom_point(aes(x=as.Date(data$date), y=data$WH))+
  xlab("")+
  ylab("Reported and adjusted cases")+
  ggtitle("Western Health")+
  #ggtitle(paste("At entry: ", exp.on.entry,"days since exposure"))+
theme_classic()+
  scale_x_date(date_breaks = "2 day", date_labels = "%d %b")+
  theme(axis.text.x = element_text(angle = 90),legend.title = element_blank())

g.LG=ggplot() +
geom_line(data=data2.corr, aes(x=as.Date(date), y=LG, group=1))+
geom_point(aes(x=as.Date(data$date), y=data$LG))+
  xlab("")+
  ylab("Reported and adjusted cases")+
  ggtitle("Labrador-Grenfell Health")+
  #ggtitle(paste("At entry: ", exp.on.entry,"days since exposure"))+
theme_classic()+
  scale_x_date(date_breaks = "2 day", date_labels = "%d %b")+
  theme(axis.text.x = element_text(angle = 90),legend.title = element_blank())

g
g.EH+g.CH
g.WH+g.LG
```

**Figure 1.** The adjustment for the PCR testing change and out-of-province reporting (line) and reported cases (dots).

```{r,echo=F}
g1
```

**Figure 2.** Data is adjusted for the changes to PCR testing and the reporting of testing performed out-of-province.

## Doubling and halving times

The doubling time is the days until the number of reported cases doubles (assuming exponential growth). The halving time is the days until the number of reported cases halves (assuming exponential decay). Our doubling and halving time calculations consider the number of reported cases over the last `r T` days. If the exponential growth coefficient is not different from 0, the number of reported cases is stated as 'level'.

Due to assumed pending test results that have been sent out-of-province for Eastern Health and Labrador, we exclude the last 3 days from calculations of doubling/halving time.

**Newfoundland and Labrador: Declining with a halving time of `r DT.NL[1]` days** 

**Eastern Health: level** 

**Central Health: Declining with a halving time of `r DT.central[1]` days** 

**Western Health: level** 

**Labrador-Grenfell Health: level**

Fitting is performed using nonlinear least squares and considers the adjusted data.

```{r, include=F, warning=FALSE}
 
# Assume 3 days hospitalized. Admissions - discharges
# Hospitalized within 14 days of specimen, take 7 as mean.
d = 3
# Cases per hospitlation/death
UKh = (198348/815)
UKm = (198348/57)
ONh = (29594/75)
ONm = (29594/3)
NLh = 1.3*ONh
NLm = 0.4*UKm
UK.rate = NL.corr/UKh
UK.rate = cumsum(UK.rate)
UK.rate[7:L] = UK.rate[7:L] - UK.rate[1:(L-6)]
ON.rate = NL.corr/ONh
ON.rate = cumsum(ON.rate)
ON.rate[7:L] = ON.rate[7:L] - ON.rate[1:(L-6)]
NL.rate = NL.corr/NLh 
NL.rate = cumsum(NL.rate)
NL.rate[7:L] = NL.rate[7:L] - NL.rate[1:(L-6)]

UK.rate.ad = cumsum(NL.corr)/UKh
ON.rate.ad = cumsum(NL.corr)/ONh
NL.rate.ad = cumsum(NL.corr)/NLh

# 649834/57 - 5 days median time to death.
UK.rate.m = cumsum(NL.corr)/UKm
ON.rate.m = cumsum(NL.corr)/ONm
NL.rate.m = cumsum(NL.corr)/NLm
future.dates = tail(seq(from=as.Date(data$date[L], format = "%Y-%m-%d"), length.out = 8, by = "days"),-1)
days2 = c(tail(as.Date(data$date, format = "%Y-%m-%d"),-7),future.dates)
future.dates.5 = tail(seq(from=as.Date(data$date[L], format = "%Y-%m-%d"), length.out = 6, by = "days"),-1)
days3 = c(tail(as.Date(data$date, format = "%Y-%m-%d"),-5),future.dates.5)
NL.hosp = c(tail(NL.data$cum.hospital,-7), rep(NA,7))
NL.mort = c(tail(NL.data$cum.mort,-5), rep(NA,5))
data2 = data.frame(dates = rep(days2,3), occupancy = c(ON.rate,UK.rate,NL.rate), location = rep(c("ON", "UK", "NL"), each = L))
data3 = data.frame(dates = rep(days3,3), mortality = c(ON.rate.m, UK.rate.m,NL.rate.m), location = rep(c("ON", "UK", "NL"), each = length(days3)))
data4 = data.frame(dates = rep(days2,3), admission = c(UK.rate.ad, ON.rate.ad, NL.rate.ad),location = rep(c("ON", "UK", "NL"), each = L)) 
pointdata = data.frame(days2 = days2, hospital = NL.hosp)
pointdata2 = data.frame(days2 = days3, mortality = NL.mort)
```

```{r, include=FALSE, warning=FALSE}
g2=ggplot()+
geom_line(data=data2, aes(x=as.Date(dates), y=occupancy, group = location, color=location))+
  geom_point(data = pointdata, aes(x = days2, y = hospital))+
  geom_line(data = pointdata, aes(x = days2, y = hospital))+
  xlab("")+
  ylab("")+
  scale_x_date(date_breaks = "1 day", date_labels = "%d %b")+
  ggtitle("Projected and actual hospital occupancy")+
theme_classic()+
  theme(axis.text.x = element_text(angle = 90),legend.title = element_blank())

g2b=ggplot()+
geom_line(data=data4, aes(x=as.Date(dates), y=admission, group = location, color=location))+
  xlab("")+
  ylab("")+
  scale_x_date(date_breaks = "1 day", date_labels = "%d %b")+
  ggtitle("Projected cumulative hospitalizations since December 15, 2021")+
theme_classic()+
  theme(axis.text.x = element_text(angle = 90),legend.title = element_blank())

g3=ggplot()+
geom_line(data=data3, aes(x=as.Date(dates), y=mortality, group = location, color=location))+
  geom_point(data = pointdata2, aes(x = days2, y = mortality))+
  geom_line(data = pointdata2, aes(x = days2, y = mortality))+
  xlab("")+
  ylab("")+
  scale_x_date(date_breaks = "1 day", date_labels = "%d %b")+
  ggtitle("Projected and actual cumulative deaths since December 15, 2021")+
theme_classic()+
  theme(axis.text.x = element_text(angle = 90),legend.title = element_blank())
```

## Hospitalizations and deaths

We consider data from [United Kingdom Health Security Agency  (UKHSA)](https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/1044481/Technical-Briefing-31-Dec-2021-Omicron_severity_update.pdf) and [Public Health Ontario (PHO)](https://www.publichealthontario.ca/-/media/documents/ncov/epi/covid-19-epi-enhanced-estimates-omicron-severity-study.pdf?sc_lang=en) reports to understand the relationship between reported Omicron cases and the future number of hospitalizations and deaths.

The UKHSA study reports for confirmed Omicron cases: 198,348 cases, 815 hospital admissions (0.41%), and 57 deaths (0.03%) up until December 29, 2021. The PHO study reports for Omicron: 29,594 cases, 75 hospitalizations or deaths (0.25%), and 3 deaths (0.01%) from November 22 - December 17, 2021.

```{r, echo=FALSE,warning=FALSE}
g2
```

**Figure 3.** The black dots and lines are the number of people in hospital as reported by the NL [Public Advisories](https://www.gov.nl.ca/releases/covid-19-news/). We assume hospitalizations are lagged 7 days from cases and the length of stay in hospital is 6 days ([BC Modelling Report, January 6, 2022](https://bccovid-19group.ca/post/2022-01-06-report/COVID19group-2022-01-06.pdf)). Hospital occupancy when assuming the reported proportion of hospitalizations for Omicron cases in the UKHSA (blue) and PHO (green) studies. The calibration to the NL data (red) assumes `r round(100/NLh,2)`% of reported cases are hospitalized. These calculations consider the adjusted case data.

<!-- ```{r, echo=FALSE,warning=FALSE} -->
<!-- g2b -->
<!-- ``` -->

<!-- **Figure 4.** The reported proportion of hospitalizations for Omicron cases in the PHO and UKHSA studies are parameters that determine the lower and upper bounds of the grey shaded regions for projected cumulative number of hospitalizations since December 15th, 2021 in NL. In addition, we assume hospitalizations are lagged 7 days from cases. The NL Public Advisories do not report the number of new admissions or discharges, so these predictions of the cumulative number of hospitalizations cannot be evaluated using the publicly available information. The range of cumulative hospitalization is [`r round(ON.rate.ad[(L-7)],2)`, `r round(UK.rate.ad[(L-7)],2)`] on `r as.Date(days2[(L-7)], format = "%d %b")` and [`r round(ON.rate.ad[L],2)`, `r round(UK.rate.ad[L],2)`] on `r as.Date(days2[L], format= "%d %b")`. -->


```{r, echo=FALSE,warning=FALSE}
g3
```

**Figure 4.** The black dots and lines are the number of deaths as reported by the NL [Public Advisories](https://www.gov.nl.ca/releases/covid-19-news/). We assume the time from being reported as a case to death is 5 days, which is the median time from specimen to death in the UKHSA study. Cumulative deaths since December 15, 2021 when assuming the reported proportion of deaths for Omicron cases in the UKHSA (blue) and PHO (green) studies. The calibration to the NL data (red) assumes that death occurs for `r round(100/NLm,2)`% of reported cases. These calculations consider the adjusted case data.


